---
title: "Chicken_Pox"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(keras)
```

## load Data
```{r}
dat <- read_csv("hungary_chickenpox/hungary_chickenpox.csv")
head(dat)
dim(dat)
```


##Exercises:

### 0a. Do you want to center and scale?  If so don't forget to do it.

I feel like I should given the different ranges of each county

### 0b. Is it fair to use week as a predictor? IT is debatable, but think about it and act appropriately.  You might try it both ways.  I am going to start without.

I wouldn't since the week used changes from year to year. A better predictor would probably be month

1. plot the data in an informative way

```{r}
dat %>%
  pivot_longer(2:21, names_to = "County", values_to = "Cases") %>%
  ggplot(aes(x = Date, y = Cases)) +
  geom_line()
```


### 2. Because this is a small data set and we are just practicing we are not going to set up a test set, just training and validation.  Set up your generators or data so that training data is 2005-2012 and validation is 2013, 2014.

```{r}
# Set train data set
data <- data.matrix(dat[,-1])
train_data <- data[1:418,] # first 8 years
mean <- apply(train_data, 2, mean)
std <- apply(train_data, 2, sd)
data <- scale(data, center = mean, scale = std)
```

lookback = How far to go back, each timestep is 1 week
steps = How many steps to take, each step is 1 week
delay = How far ahead to predict, each step is 1 week


data = The original array of floating-point data
lookback = How many timesteps back the input data should go.
delay = How many timesteps in the future the target should be.
min_index and max_index = Indices in the data array that delimit which timesteps to draw from. This is useful for keeping a segment of the data for validation and another for testing.
shuffle = Whether to shuffle the samples or draw them in chronological order.
batch_size = The number of samples per batch.
step = The period, in timesteps, at which you sample data

```{r}
#Make data long?
data_long <- data %>%
  data.frame() %>%
  pivot_longer(BUDAPEST:ZALA, names_to = "County", values_to = "Cases")
data_long_matrix <- data.matrix(data_long[,-1])
(418 * 20) # first 8360 rows are the first 8 years
```


```{r}
#Build Generator
generator <- function(data, lookback, delay, min_index, max_index,
                      shuffle = FALSE, batch_size = 128, step = 6) {
  if (is.null(max_index)) max_index <- nrow(data) - delay - 1
  i <- min_index + lookback
  function() {
    if (shuffle) {
      rows <- sample(c((min_index+lookback):max_index), size = batch_size)
    } else {
      if (i + batch_size >= max_index)
        i <<- min_index + lookback
      rows <- c(i:min(i+batch_size, max_index))
      i <<- i + length(rows)
    }
    samples <- array(0, dim = c(length(rows),
                                lookback / step,
                                dim(data)[[-1]]))
    targets <- array(0, dim = c(length(rows)))
    for (j in 1:length(rows)) {
      indices <- seq(rows[[j]] - lookback, rows[[j]],
                     length.out = dim(samples)[[2]])
      samples[j,,] <- data[indices,]
      targets[[j]] <- data[rows[[j]] + delay,1]
    }
    list(samples, targets)
  }
}
```

```{r}
lookback <- 1
step <- 1
delay <- 1
batch_size <- 32
train_gen <- generator(
  data_long_matrix,
  lookback = lookback,
  delay = delay,
  min_index = 1,
  max_index = 8360,
  shuffle = TRUE,
  step = step,
  batch_size = batch_size
)
val_gen = generator(
  data_long_matrix,
  lookback = lookback,
  delay = delay,
  min_index = 8361,
  max_index = 10440,
  step = step,
  batch_size = batch_size
)
```

```{r}
val_steps <- (10440 - 8361 - lookback) / batch_size
```

```{r}
# baseline
evaluate_naive_method <- function() {
  batch_maes <- c()
  for (step in 1:val_steps) {
    c(samples, targets) %<-% val_gen()
    preds <- samples[,dim(samples)[[2]],1]
    mae <- mean(abs(preds - targets))
    batch_maes <- c(batch_maes, mae)
  }
  print(mean(batch_maes))
}
evaluate_naive_method()
```


3. Predict chickenpox case # in Budapest 2 weeks in the future.  Be sure to include a baseline model for comparison.  

```{r}
model <- keras_model_sequential() %>%
  layer_flatten(input_shape = c(lookback / step, dim(data_long_matrix)[-1])) %>%
  layer_dense(units = 32, activation = "relu") %>%
  layer_dense(units = 1)
model %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "mae"
)
history <- model %>% fit(
  train_gen,
  steps_per_epoch = 500,
  epochs = 20,
  validation_data = val_gen,
  validation_steps = val_steps
)
```


4. Recode the data so that you can make predicition for all counties 2 weeks in the future.  Be sure to include a baseline model for comparison.

5. If you are getting good predictions, how good can you do at 1 month?  If predictions are poor maybe try going down to 1 week.