---
title: "Object_Detection"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rjson)
library(magick)
library(purrr)
library(tibble)
library(tidyr)
library(dplyr)
library(ggplot2)
library(stringr)
library(keras)
use_condaenv('r-reticulate')
library(tensorflow)
```

```{r}
physical_devices <- tf$config$list_physical_devices('GPU')
tf$config$experimental$set_memory_growth(device = physical_devices[[1]], enable = T)
```


```{r}
img_dir <- "Object_Detection/VOCdevkit/VOC2007/JPEGImages"
annot_file <- "Object_Detection/pascal_train2007.json"
```

```{r}
annotations <- fromJSON(file = annot_file)
str(annotations, max.level = 1)
```

```{r}
imageinfo <- annotations$images %>% {
  tibble(
    id = map_dbl(., "id"),
    file_name = map_chr(., "file_name"),
    image_height = map_dbl(., "height"),
    image_width = map_dbl(., "width")
  )
}
```

```{r}
classes <- c(
  "aeroplane",
  "bicycle",
  "bird",
  "boat",
  "bottle",
  "bus",
  "car",
  "cat",
  "chair",
  "cow",
  "diningtable",
  "dog",
  "horse",
  "motorbike",
  "person",
  "pottedplant",
  "sheep",
  "sofa",
  "train",
  "tvmonitor"
)
```

```{r}
boxinfo <- annotations$annotations %>% {
  tibble(
    image_id = map_dbl(., "image_id"),
    category_id = map_dbl(., "category_id"),
    bbox = map(., "bbox")
  )
}
```

```{r}
boxinfo <- boxinfo %>% 
  mutate(bbox = unlist(map(.$bbox, function(x) paste(x, collapse = " "))))
boxinfo <- boxinfo %>% 
  separate(bbox, into = c("x_left", "y_top", "bbox_width", "bbox_height"))
boxinfo <- boxinfo %>% mutate_all(as.numeric)
```

```{r}
boxinfo <- boxinfo %>% 
  mutate(y_bottom = y_top + bbox_height - 1, x_right = x_left + bbox_width - 1)
```

```{r}
catinfo <- annotations$categories %>%  {
  tibble(id = map_dbl(., "id"), name = map_chr(., "name"))
}
```

```{r}
imageinfo <- imageinfo %>%
  inner_join(boxinfo, by = c("id" = "image_id")) %>%
  inner_join(catinfo, by = c("category_id" = "id"))
```

```{r}
target_height <- 224
target_width <- 224

imageinfo <- imageinfo %>% mutate(
  x_left_scaled = (x_left / image_width * target_width) %>% round(),
  x_right_scaled = (x_right / image_width * target_width) %>% round(),
  y_top_scaled = (y_top / image_height * target_height) %>% round(),
  y_bottom_scaled = (y_bottom / image_height * target_height) %>% round(),
  bbox_width_scaled =  (bbox_width / image_width * target_width) %>% round(),
  bbox_height_scaled = (bbox_height / image_height * target_height) %>% round()
)
```

```{r}
img_data <- imageinfo[4,]
img <- image_read(file.path(img_dir, img_data$file_name))
img <- image_draw(img)
rect(
  img_data$x_left,
  img_data$y_bottom,
  img_data$x_right,
  img_data$y_top,
  border = "purple",
  lwd = 2
)
text(
  img_data$x_right,
  img_data$y_top,
  img_data$name,
  offset = 1,
  pos = 2,
  cex = 1.5,
  col = "purple"
)
dev.off()
```

```{r}
imageinfo <- imageinfo %>% mutate(area = bbox_width_scaled * bbox_height_scaled)

imageinfo_maxbb <- imageinfo %>%
  group_by(id) %>%
  filter(which.max(area) == row_number())
```

```{r}
n_samples <- nrow(imageinfo_maxbb)
train_indices <- sample(1:n_samples, 0.8 * n_samples)
train_data <- imageinfo_maxbb[train_indices,]
validation_data <- imageinfo_maxbb[-train_indices,]
```

```{r}
feature_extractor <-
  application_xception(
    include_top = FALSE,
    input_shape = c(224, 224, 3),
    pooling = "avg"
)

feature_extractor %>% freeze_weights()
```

```{r}
model <- keras_model_sequential() %>%
  feature_extractor %>%
  layer_batch_normalization() %>%
  layer_dropout(rate = 0.25) %>%
  layer_dense(units = 512, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 20, activation = "softmax")

model %>% compile(
  optimizer = "adam",
  loss = "sparse_categorical_crossentropy",
  metrics = list("accuracy")
)
```

```{r}
summary(model)
```

```{r}
batch_size <- 10

load_and_preprocess_image <- function(image_name, target_height, target_width) {
  img_array <- image_load(
    file.path(img_dir, image_name),
    target_size = c(target_height, target_width)
    ) %>%
    image_to_array() %>%
    xception_preprocess_input() 
  dim(img_array) <- c(1, dim(img_array))
  img_array
}

classification_generator <-
  function(data,
           target_height,
           target_width,
           shuffle,
           batch_size) {
    i <- 1
    function() {
      if (shuffle) {
        indices <- sample(1:nrow(data), size = batch_size)
      } else {
        if (i + batch_size >= nrow(data))
          i <<- 1
        indices <- c(i:min(i + batch_size - 1, nrow(data)))
        i <<- i + length(indices)
      }
      x <-
        array(0, dim = c(length(indices), target_height, target_width, 3))
      y <- array(0, dim = c(length(indices), 1))
      
      for (j in 1:length(indices)) {
        x[j, , , ] <-
          load_and_preprocess_image(data[[indices[j], "file_name"]],
                                    target_height, target_width)
        y[j, ] <-
          data[[indices[j], "category_id"]] - 1
      }
      x <- x / 255
      list(x, y)
    }
  }

train_gen <- classification_generator(
  train_data,
  target_height = target_height,
  target_width = target_width,
  shuffle = TRUE,
  batch_size = batch_size
)

valid_gen <- classification_generator(
  validation_data,
  target_height = target_height,
  target_width = target_width,
  shuffle = FALSE,
  batch_size = batch_size
)
```

```{r}
model %>% fit(
  train_gen,
  epochs = 9,
  steps_per_epoch = nrow(train_data) / batch_size,
  validation_data = valid_gen,
  validation_steps = nrow(validation_data) / batch_size,
  callbacks = list(
    callback_model_checkpoint(
      file.path("class_only", "weights.{epoch:02d}-{val_loss:.2f}.hdf5")
    ),
    callback_early_stopping(patience = 2)
  )
)
```
# Died after 9 epochs first time

```{r}
image_cats <- imageinfo %>% 
  select(category_id) %>%
  mutate(category_id = category_id - 1) %>%
  pull() %>%
  to_categorical(num_classes = 20)

image_cats <- data.frame(image_cats) %>%
  add_column(file_name = imageinfo$file_name, .before = TRUE)

image_cats <- image_cats %>% 
  group_by(file_name) %>% 
  summarise_all(.funs = funs(max))

n_samples <- nrow(image_cats)
train_indices <- sample(1:n_samples, 0.8 * n_samples)
train_data <- image_cats[train_indices,]
validation_data <- image_cats[-train_indices,]
```

```{r}
classification_generator <- 
  function(data,
           target_height,
           target_width,
           shuffle,
           batch_size) {
    i <- 1
    function() {
      if (shuffle) {
        indices <- sample(1:nrow(data), size = batch_size)
      } else {
        if (i + batch_size >= nrow(data))
          i <<- 1
        indices <- c(i:min(i + batch_size - 1, nrow(data)))
        i <<- i + length(indices)
      }
      x <-
        array(0, dim = c(length(indices), target_height, target_width, 3))
      y <- array(0, dim = c(length(indices), 20))
      
      for (j in 1:length(indices)) {
        x[j, , , ] <-
          load_and_preprocess_image(data[[indices[j], "file_name"]], 
                                    target_height, target_width)
        y[j, ] <-
          data[indices[j], 2:21] %>% as.matrix()
      }
      x <- x / 255
      list(x, y)
    }
  }

train_gen <- classification_generator(
  train_data,
  target_height = target_height,
  target_width = target_width,
  shuffle = TRUE,
  batch_size = batch_size
)

valid_gen <- classification_generator(
  validation_data,
  target_height = target_height,
  target_width = target_width,
  shuffle = FALSE,
  batch_size = batch_size
)
```

```{r}
feature_extractor <-
  application_xception(
    include_top = FALSE,
    input_shape = c(224, 224, 3),
    pooling = "avg"
  )

feature_extractor %>% freeze_weights()

model <- keras_model_sequential() %>%
  feature_extractor %>%
  layer_batch_normalization() %>%
  layer_dropout(rate = 0.25) %>%
  layer_dense(units = 512, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 20, activation = "sigmoid")

model %>% compile(optimizer = "adam",
                  loss = "binary_crossentropy",
                  metrics = list("accuracy"))
```

Adam optimization is a stochastic gradient descent method that is based on adaptive estimation of first-order and second-order moments.

According to Kingma et al., 2014, the method is "computationally efficient, has little memory requirement, invariant to diagonal rescaling of gradients, and is well suited for problems that are large in terms of data/parameters".

```{r}
model %>% fit(
  train_gen,
  epochs = 20,
  steps_per_epoch = nrow(train_data) / batch_size,
  validation_data = valid_gen,
  validation_steps = nrow(validation_data) / batch_size,
  callbacks = list(
    callback_model_checkpoint(
      file.path("multiclass", "weights.{epoch:02d}-{val_loss:.2f}.hdf5")
    ),
    callback_early_stopping(patience = 2)
  )
)
```
Stopped after 5 epochs

```{r}
feature_extractor <- application_xception(
  include_top = FALSE,
  input_shape = c(224, 224, 3)
)

feature_extractor %>% freeze_weights()
```

```{r}
model <- keras_model_sequential() %>%
  feature_extractor %>%
  layer_flatten() %>%
  layer_batch_normalization() %>%
  layer_dropout(rate = 0.25) %>%
  layer_dense(units = 512, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 4)
```

```{r}
metric_iou <- function(y_true, y_pred) {
  
  # order is [x_left, y_top, x_right, y_bottom]
  intersection_xmin <- k_maximum(y_true[ ,1], y_pred[ ,1])
  intersection_ymin <- k_maximum(y_true[ ,2], y_pred[ ,2])
  intersection_xmax <- k_minimum(y_true[ ,3], y_pred[ ,3])
  intersection_ymax <- k_minimum(y_true[ ,4], y_pred[ ,4])
  
  area_intersection <- (intersection_xmax - intersection_xmin) * 
                       (intersection_ymax - intersection_ymin)
  area_y <- (y_true[ ,3] - y_true[ ,1]) * (y_true[ ,4] - y_true[ ,2])
  area_yhat <- (y_pred[ ,3] - y_pred[ ,1]) * (y_pred[ ,4] - y_pred[ ,2])
  area_union <- area_y + area_yhat - area_intersection
  
  iou <- area_intersection/area_union
  k_mean(iou)
  
}
```

```{r}
model %>% compile(
  optimizer = "adam",
  loss = "mae",
  metrics = list(custom_metric("iou", metric_iou))
)
```

# Need to fix train data since not using categories

```{r}
n_samples <- nrow(imageinfo_maxbb)
train_indices <- sample(1:n_samples, 0.8 * n_samples)
train_data <- imageinfo_maxbb[train_indices,]
validation_data <- imageinfo_maxbb[-train_indices,]
```


```{r}
localization_generator <-
  function(data,
           target_height,
           target_width,
           shuffle,
           batch_size) {
    i <- 1
    function() {
      if (shuffle) {
        indices <- sample(1:nrow(data), size = batch_size)
      } else {
        if (i + batch_size >= nrow(data))
          i <<- 1
        indices <- c(i:min(i + batch_size - 1, nrow(data)))
        i <<- i + length(indices)
      }
      x <-
        array(0, dim = c(length(indices), target_height, target_width, 3))
      y <- array(0, dim = c(length(indices), 4))
      
      for (j in 1:length(indices)) {
        x[j, , , ] <-
          load_and_preprocess_image(data[[indices[j], "file_name"]], 
                                    target_height, target_width)
        y[j, ] <-
          data[indices[j], c("x_left_scaled",
                             "y_top_scaled",
                             "x_right_scaled",
                             "y_bottom_scaled")] %>% as.matrix()
      }
      x <- x / 255
      list(x, y)
    }
  }

train_gen <- localization_generator(
  train_data,
  target_height = target_height,
  target_width = target_width,
  shuffle = TRUE,
  batch_size = batch_size
)

valid_gen <- localization_generator(
  validation_data,
  target_height = target_height,
  target_width = target_width,
  shuffle = FALSE,
  batch_size = batch_size
)
```


```{r}
model %>% fit(
  train_gen,
  epochs = 20,
  steps_per_epoch = nrow(train_data) / batch_size,
  validation_data = valid_gen,
  validation_steps = nrow(validation_data) / batch_size,
  callbacks = list(
    callback_model_checkpoint(
      file.path("loc_only", "weights.{epoch:02d}-{val_loss:.2f}.hdf5")
    ),
    callback_early_stopping(patience = 2)
  )
)
```

```{r}
plot_image_with_boxes <- function(file_name,
                                  object_class,
                                  box,
                                  scaled = FALSE,
                                  class_pred = NULL,
                                  box_pred = NULL) {
  img <- image_read(file.path(img_dir, file_name))
  if(scaled) img <- image_resize(img, geometry = "224x224!")
  img <- image_draw(img)
  x_left <- box[1]
  y_bottom <- box[2]
  x_right <- box[3]
  y_top <- box[4]
  rect(
    x_left,
    y_bottom,
    x_right,
    y_top,
    border = "cyan",
    lwd = 2.5
  )
  text(
    x_left,
    y_top,
    object_class,
    offset = 1,
    pos = 2,
    cex = 1.5,
    col = "cyan"
  )
  if (!is.null(box_pred)){
    rect(box_pred[1],
         box_pred[2],
         box_pred[3],
         box_pred[4],
         border = "yellow",
         lwd = 2.5)
  }
  if (!is.null(class_pred)){
    text(
      box_pred[1],
      box_pred[2],
      class_pred,
      offset = 0,
      pos = 4,
      cex = 1.5,
      col = "yellow")
  }
  dev.off()
  img %>% image_write(paste0("preds_", file_name))
  plot(img)
}
```

```{r}
train_1_8 <- train_data[1:8, c("file_name",
                               "name",
                               "x_left_scaled",
                               "y_top_scaled",
                               "x_right_scaled",
                               "y_bottom_scaled")]

for (i in 1:8) {
  preds <-
    model %>% predict(
      load_and_preprocess_image(train_1_8[i, "file_name"], 
                                target_height, target_width),
      batch_size = 1
  )
  print(preds)
  plot_image_with_boxes(train_1_8$file_name[i],
                        train_1_8$name[i],
                        train_1_8[i, 3:6] %>% as.matrix(),
                        scaled = TRUE,
                        box_pred = preds)
}
```

#Predictions too bad to show up

```{r}
validation_1_8 <- validation_data[1:8, c("file_name",
                               "name",
                               "x_left_scaled",
                               "y_top_scaled",
                               "x_right_scaled",
                               "y_bottom_scaled")]

for (i in 1:8) {
  preds <-
    model %>% predict(
      load_and_preprocess_image(validation_1_8[i, "file_name"], 
                                target_height, target_width),
      batch_size = 1
  )
  print(preds)
  plot_image_with_boxes(validation_1_8$file_name[i],
                        validation_1_8$name[i],
                        validation_1_8[i, 3:6] %>% as.matrix(),
                        scaled = TRUE,
                        box_pred = preds)
}
```

trash

```{r}
feature_extractor <- application_xception(
  include_top = FALSE,
  input_shape = c(224, 224, 3)
)

input <- feature_extractor$input
common <- feature_extractor$output %>%
  layer_flatten(name = "flatten") %>%
  layer_activation_relu() %>%
  layer_dropout(rate = 0.25) %>%
  layer_dense(units = 512, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dropout(rate = 0.5)

regression_output <-
  layer_dense(common, units = 4, name = "regression_output")
class_output <- layer_dense(
  common,
  units = 20,
  activation = "softmax",
  name = "class_output"
)

model <- keras_model(
  inputs = input,
  outputs = list(regression_output, class_output)
)
```

```{r}
model %>% freeze_weights(to = "flatten")

model %>% compile(
  optimizer = "adam",
  loss = list("mae", "sparse_categorical_crossentropy"),
  #loss_weights = list(
  #  regression_output = 0.05,
  #  class_output = 0.95),
  metrics = list(
    regression_output = custom_metric("iou", metric_iou),
    class_output = "accuracy"
  )
)
```

```{r}
loc_class_generator <-
  function(data,
           target_height,
           target_width,
           shuffle,
           batch_size) {
    i <- 1
    function() {
      if (shuffle) {
        indices <- sample(1:nrow(data), size = batch_size)
      } else {
        if (i + batch_size >= nrow(data))
          i <<- 1
        indices <- c(i:min(i + batch_size - 1, nrow(data)))
        i <<- i + length(indices)
      }
      x <-
        array(0, dim = c(length(indices), target_height, target_width, 3))
      y1 <- array(0, dim = c(length(indices), 4))
      y2 <- array(0, dim = c(length(indices), 1))
      
      for (j in 1:length(indices)) {
        x[j, , , ] <-
          load_and_preprocess_image(data[[indices[j], "file_name"]], 
                                    target_height, target_width)
        y1[j, ] <-
          data[indices[j], c("x_left", "y_top", "x_right", "y_bottom")] %>% 
          as.matrix()
        y2[j, ] <-
          data[[indices[j], "category_id"]] - 1
      }
      x <- x / 255
      list(x, list(y1, y2))
    }
  }

train_gen <- loc_class_generator(
  train_data,
  target_height = target_height,
  target_width = target_width,
  shuffle = TRUE,
  batch_size = batch_size
)

valid_gen <- loc_class_generator(
  validation_data,
  target_height = target_height,
  target_width = target_width,
  shuffle = FALSE,
  batch_size = batch_size
)

model %>% fit(
  train_gen,
  epochs = 20,
  steps_per_epoch = nrow(train_data) / batch_size,
  validation_data = valid_gen,
  validation_steps = nrow(validation_data) / batch_size,
  callbacks = list(
    callback_model_checkpoint(
      file.path("loc_class", "weights.{epoch:02d}-{val_loss:.2f}.hdf5")
    ),
    callback_early_stopping(patience = 2)
  )
)
```

```{r}
imageinfo %>% group_by(name) %>% 
  summarise(cnt = n()) %>%
  arrange(desc(cnt))
```

